---
title: "p8105_hw3_ls3751"
author: "Liucheng Shi"
output: github_document
---

_packages required_
```{r setup, message = F}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and ... columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```


Let's make a plot

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Let's make a table!!

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```

## Problem 2

#### 1.1.1 load the dataset and tidy data

```{r}
accel_df = 
  read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "minute",
    names_prefix = "activity_", 
    values_to = "counts"
  )
```

### 1.1.2 Create a variable that indicate the weekday/weekend status and modify variable formats

```{r}
accel_df = accel_df %>% 
  mutate(
    weekday_weekend = case_when(
      day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",
      day %in% c("Saturday", "Sunday") ~ "weekend",
      TRUE ~ "")
      ) %>% 
  mutate(
    minute = as.numeric(minute),
    week = factor(week),
    day = forcats::fct_relevel(day, c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
    )) %>% 
  relocate(weekday_weekend, .before = minute)
```

After data cleaning and wrangling, this dataset has `r nrow(accel_df)` observations with `r ncol(accel_df)` variables, including `r names(accel_df)`. 
*   _week_id_: the week indicator ranging from 1 to 5
*   _day_id_: the day indicator ranging from 1 to 35 (5 weeks in total)
*   _day_ & _weekday_weekend_: ex: Sunday, weekend 
*   _minutes_: the recorded time  (1440 min a day) 
*   _counts_: the number of activities measured by accelerometer. 

### 1.2

```{r}
accel_df %>% 
  group_by(day_id) %>% 
  summarise(n_each = sum(counts))
```

